---
title: "4.5 Model Comparison: RMSE and Final Selection"
output:
  pdf_document: default
  html_document: default
date: "2026-04-09"
---

## Goal

This section will compare forecasting performance across all methods using holdout RMSE and select the final model.

Methods included:

- Simple Rules (baseline model)
- Exponential Smoothing (ETS)
- ARIMA
- ARMAX

The primary evaluation metric is holdout RMSE, which reflects out-of-sample predictive accuracy.

## RMSE Caluclations
The following computations complete RMSE calculations for finalized models derived from previous analysis. 


#### RMSE for ARIMA
First for all the candidate ARIMA models:

```{r}
library(tidyverse)
library(forecast)

agg <- readr::read_csv(
  "../data/processed/store_type_a_train_aggregate_daily.csv"
) |>
  dplyr::mutate(Date = as.Date(Date)) |>
  dplyr::arrange(Date)

sales_ts <- ts(agg$Sales, frequency = 7)

holdout_prop <- 0.20
n_total <- nrow(agg)
n_train <- floor((1 - holdout_prop) * n_total)

train_subset <- agg |> slice(1:n_train)
holdout_subset <- agg |> slice((n_train + 1):n_total)

y_all <- ts(agg$Sales, frequency = 7)

model1 <- auto.arima(ts(train_subset$Sales, frequency = 7))

model2 <- Arima(ts(train_subset$Sales, frequency = 7), order=c(4,0,1), seasonal=c(2,1,1))
model3 <- Arima(ts(train_subset$Sales, frequency = 7), order=c(4,0,1), seasonal=c(1,1,0))
model4 <- Arima(ts(train_subset$Sales, frequency = 7), order=c(2,0,1), seasonal=c(2,1,0))
model5 <- Arima(ts(train_subset$Sales, frequency = 7), order=c(4,0,2), seasonal=c(2,1,0))

models <- list(
  ARIMA_auto = model1,
  ARIMA_m2 = model2,
  ARIMA_m3 = model3,
  ARIMA_m4 = model4,
  ARIMA_m5 = model5
)

rmse_vec <- function(actual, pred) {
  sqrt(mean((actual - pred)^2, na.rm = TRUE))
}

# RMSE computation using 1-step forecast
arima_rmse_table <- purrr::imap_dfr(models, function(mod, name) {

  ord <- arimaorder(mod)

  fixed_fit <- stats::arima(
    y_all,
    order = c(ord["p"], ord["d"], ord["q"]),
    seasonal = list(order = c(ord["P"], ord["D"], ord["Q"]), period = ord["m"]),
    fixed = as.numeric(mod$coef),
    transform.pars = FALSE,
    optim.control = list(maxit = 0),
    method = "ML"
  )

  fc_all <- as.numeric(y_all - fixed_fit$residuals)

  holdout_idx <- (n_train + 1):length(y_all)

  preds <- tibble::tibble(
    model = name,
    rmse = rmse_vec(
      actual = as.numeric(y_all[holdout_idx]),
      pred = fc_all[holdout_idx]
    )
  )

  preds
}) |>
  arrange(rmse)

arima_rmse_table
```
#### Interpretation:
Based on the results above, the best ARIMA model out of all candidates is model with RMSE = . We will proceed with this model as a representative for the ARIMA method in the remaining analysis.


#### RMSE for ARMAX
Then for ARMAX models:


## Combining RMSE Results Across Explored Methods

```{r}
simple_rmse <- readr::read_csv(
  "../04_analysis/simple_rules/results/simple_rules_aggregate_holdout_rmse_summary.csv",
  show_col_types = FALSE
) |>
  dplyr::mutate(method = "Simple")

ets_rmse <- readr::read_csv(
  "../04_analysis/ets/results/ets_holdout_rmse_summary.csv",
  show_col_types = FALSE
) |>
  dplyr::mutate(method = "ETS")


arima_rmse <- tibble(
 model = "ARIMA_best",
 holdout_rmse = <value>,
 n_holdout = <value>,
 method = "ARIMA"
)

# ARMAX (TEMP placeholder — will be added after your section)
# armax_rmse <- readr::read_csv("...") |> mutate(method = "ARMAX")

rmse_all <- dplyr::bind_rows(
  simple_rmse,
  ets_rmse
  arima_rmse,
  armax_rmse
)

rmse_all <- rmse_all |>
  dplyr::mutate(
    model_full = paste(method, model, sep = " - ")
  ) |>
  dplyr::arrange(holdout_rmse)

rmse_all
```

#### Interpretation

The table above compares forecasting performance across all finalized models, where:

- Lower RMSE indicates better predictive accuracy.  
- Baseline models provide a benchmark for evaluating more complex approaches.  
- Improvements over simple rules indicate that the more complex models successfully capture additional structure in the data.

Let's visualize the comparisons.
```{r}
ggplot(rmse_all, aes(x = reorder(model_full, holdout_rmse), y = holdout_rmse)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Holdout RMSE Comparison Across Models",
    x = "Model",
    y = "RMSE"
  )
```
#### Interpretation

The bar chart above highlights differences in forecasting performance:

- Models with substantially lower RMSE provide meaningful insights to how the target variable may be predicted.  
- If complex models (ARIMA/ARMAX) only marginally outperform simple rules, their added complexity may not be justified.  
- Large improvements suggest that time series structure and/or external variables are important.


```{r}
```


