---
title: "4.5 Model Comparison: RMSE and Final Selection"
output:
  pdf_document: default
  html_document: default
date: "2026-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

## Goal

This section will compare forecasting performance across all methods using holdout RMSE and select the final model.

Methods included:

- Simple Rules (baseline models)
- Exponential Smoothing (ETS)
- ARIMA models
- ARMAX models

The primary evaluation metric is holdout RMSE, which reflects out-of-sample predictive accuracy.

## RMSE Calculations
The following computations complete RMSE calculations for models derived from previous analysis. We will apply RMSE calculations to all candidate forecasting models to decide on a best-performing model with smallest RMSE.

**Step 1**: Setting up
```{r}
suppressPackageStartupMessages({
  suppressWarnings(library(tidyverse))
  suppressWarnings(library(forecast))
})

agg <- readr::read_csv(
  "../../data/processed/store_type_a_train_aggregate_daily.csv",
  show_col_types = FALSE
) |>
  dplyr::mutate(Date = as.Date(Date)) |>
  dplyr::arrange(Date)

y_all <- agg$Sales

n_total <- length(y_all)
n_train <- floor(0.8 * n_total)

train_data <- y_all[1:n_train]
test_data  <- y_all[(n_train + 1):n_total]
holdout_idx <- (n_train + 1):n_total

rmse_vec <- function(a, p) sqrt(mean((a - p)^2, na.rm = TRUE))
```

**Step 2**: Pulling Simple and ETS Models
```{r}
simple_rmse <- readr::read_csv(
  "../../04_analysis/simple/results/simple_rules_aggregate_holdout_rmse_summary.csv"
) |>
  mutate(method = "Simple")

sales_ts <- ts(train_data, frequency = 7)
ets_model <- ets(sales_ts)
ets_fc <- forecast(ets_model, h = length(test_data))$mean

ets_rmse <- tibble(
  model = "ETS",
  holdout_rmse = rmse_vec(test_data, ets_fc),
  n_holdout = length(test_data),
  method = "ETS"
)
```

**Step 3**: ARIMA with Candidate Model Evaluation
The ARIMA candidates below were selected to complement `auto.arima` with nearby low-order seasonal specifications informed by weekly seasonality (frequency = 7), ACF/PACF patterns from earlier diagnostics, and AIC-oriented model comparison.

```{r}
model1 <- auto.arima(sales_ts)
model2 <- Arima(sales_ts, order=c(4,0,1), seasonal=c(2,1,1))
model3 <- Arima(sales_ts, order=c(4,0,1), seasonal=c(1,1,0))
model4 <- Arima(sales_ts, order=c(2,0,1), seasonal=c(2,1,0))
model5 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(2,1,0))

models <- list(
  ARIMA_auto = model1,
  ARIMA_m2 = model2,
  ARIMA_m3 = model3,
  ARIMA_m4 = model4,
  ARIMA_m5 = model5
)

arima_rmse_table <- purrr::imap_dfr(models, function(mod, name) {

  fc <- forecast(mod, h = length(test_data))$mean

  tibble(
    model = name,
    holdout_rmse = rmse_vec(test_data, fc),
    n_holdout = length(test_data),
    method = "ARIMA"
  )
}) |>
  arrange(holdout_rmse)

arima_best <- arima_rmse_table |> slice(1)
arima_best
```

### ARIMA Model Finalization
Based on the results above, the best ARIMA model out of all candidates is model 2 with RMSE = 1016497 (sale value). We will proceed with this model as a representative for the ARIMA method in the remaining analysis.

**Step 4**: ARMAX
```{r}
armax_rmse <- readr::read_csv(
  "../../04_analysis/armax/results/armax_holdout_rmse_summary.csv",
  show_col_types = FALSE
)

armax_rmse <- armax_rmse |>
  mutate(method = "ARMAX")
```

## Combining RMSE Results Across Explored Methods

```{r}
rmse_all <- bind_rows(
  simple_rmse,
  ets_rmse,
  arima_best,
  armax_rmse
) |>
  mutate(model_full = paste(method, model, sep = " - ")) |>
  arrange(holdout_rmse)

rmse_all
```

### Interpretation: RMSE Table
Overall the table above compares forecasting performance across all finalized models, where:

- Lower RMSE indicates better predictive accuracy.  
- Baseline models such as persistence provide a benchmark for evaluating more complex approaches.  
- Improvements over simple rules indicate that the more complex models successfully capture additional structure in the data.

Let's visualize these comparisons.
```{r}
ggplot(rmse_all, aes(x = reorder(model_full, holdout_rmse), y = holdout_rmse)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Holdout RMSE Comparison (All Models)",
    x = "Model",
    y = "RMSE"
  )
```

**Interpretation (RMSE Plot)**

The bar chart above highlights differences in forecasting performance:

- ARMAX substantially outperforms all other methods by a wide margin, making it the clear best-performing model in this comparison.
- The large ARMAX gain is consistent with explanatory regressors (PromoShare, OpenStores, DayOfWeek, and Month) capturing demand drivers that univariate models cannot represent directly.
- ARIMA improves over most simple baselines but remains close to `iid_by_period`, indicating that day-of-week averaging alone already captures much of the weekly structure.
- The strong `iid_by_period` baseline result is plausible because averaging by day-of-week already captures much of the dominant weekly demand pattern, leaving limited additional gain for univariate ARIMA.

## Final Model Selection
```{r}
best_model <- rmse_all |> slice_min(holdout_rmse, n = 1)
best_model
```
### Interpretation: Final Model Choice
The selected model is the one with the lowest holdout RMSE (ARMAX model with RMSE = 399947.5) meaning the model demonstrates the best out-of-sample forecasting performance among the candidates.

## Final Test Forecast
```{r}
seasonal_frequency <- 7L
final_train <- readr::read_csv(
  "../../data/processed/store_type_a_train_aggregate_daily.csv",
  show_col_types = FALSE
) |>
  mutate(
    Date = as.Date(Date),
    PromoShare = if_else(OpenStores > 0, PromoStores / OpenStores, 0),
    DayOfWeek = factor(DayOfWeek, levels = 1:7),
    Month = factor(Month, levels = 1:12)
  ) |>
  arrange(Date)

final_test <- readr::read_csv(
  "../../data/processed/store_type_a_test_aggregate_daily.csv",
  show_col_types = FALSE
) |>
  mutate(
    Date = as.Date(Date),
    PromoShare = if_else(OpenStores > 0, PromoStores / OpenStores, 0),
    DayOfWeek = factor(DayOfWeek, levels = 1:7),
    Month = factor(Month, levels = 1:12)
  ) |>
  arrange(Date)

colnames(final_test)

xreg_train <- model.matrix(
  ~ PromoShare + OpenStores + DayOfWeek + Month,
  data = final_train
)[, -1]

xreg_test <- model.matrix(
  ~ PromoShare + OpenStores + DayOfWeek + Month,
  data = final_test
)[, -1]

y_train_full <- ts(final_train$Sales, frequency = seasonal_frequency)

final_armax_model <- auto.arima(
  y_train_full,
  xreg = xreg_train,
  seasonal = TRUE,
  allowmean = FALSE,
  allowdrift = FALSE,
  stepwise = TRUE,
  approximation = TRUE
)

h <- nrow(final_test)

final_forecast <- forecast(
  final_armax_model,
  xreg = xreg_test,
  h = h
)$mean

# Enforce non-negative sales
final_forecast <- pmax(0, as.numeric(final_forecast))

test_target_col <- intersect(c("Sales", "sales"), colnames(final_test))

if (length(test_target_col) == 1) {
  final_results <- tibble(
    Date = final_test$Date,
    actual = final_test[[test_target_col]],
    predicted = final_forecast
  )

  final_test_rmse <- sqrt(mean(
    (final_results$actual - final_results$predicted)^2,
    na.rm = TRUE
  ))
} else {
  final_results <- tibble(
    Date = final_test$Date,
    predicted = final_forecast
  )

  final_test_rmse <- NA_real_
  message("No actual sales column found in final_test; test RMSE cannot be computed from this file.")
}

head(final_results)

final_test_rmse
```
### Interpretation: Test Forecast
The final model is used to generate forecasts for a completely unseen test feature set. In the current aggregate test file, actual sales are not available, so a true test RMSE cannot be computed here. Holdout RMSE remains the valid basis for model comparison in this notebook; test RMSE requires a labeled test file with actual sales.
