---
title: "Exploratory Data Analysis: Store Type A"
output:
  pdf_document: default
  html_document: default
date: "2026-04-08"
---

## EDA Goals

We use this EDA to answer four modeling questions:

1. What does sales look like over time (stable, trend, or volatile)?
2. Is there repeating seasonality (weekly/monthly)?
3. Which factors affect sales (Promo, CompetitionDistance, DayOfWeek)?
4. Is the series stationary (to guide ARIMA differencing)?

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r libraries}
library(tidyverse)
library(lubridate)
library(tseries)
library(forecast)
```

## Load Data

```{r load-data}
agg <- read_csv("../data/processed/store_type_a_train_aggregate_daily.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date)

glimpse(agg)
```

## Aggregate-Daily Summary

```{r aggregate-summary}
summary(agg$Sales)
```

## Date Continuity Check

```{r date-continuity}
date_span <- as.integer(max(agg$Date) - min(agg$Date)) + 1
missing_days <- date_span - nrow(agg)

tibble(
  first_date = min(agg$Date),
  last_date = max(agg$Date),
  expected_days = date_span,
  observed_days = nrow(agg),
  missing_days = missing_days
)
```

## What Does Sales Look Like Over Time?

```{r plot-daily-sales, fig.width=11, fig.height=4}
ggplot(agg, aes(x = Date, y = Sales)) +
  geom_line() +
  labs(
    title = "Total Daily Sales - Store Type A",
    x = "Date",
    y = "Sales"
  )
```

## Promo Effect: Promo vs Non-Promo Days

```{r promo-effect}
promo_summary <- agg %>%
  mutate(PromoDay = if_else(PromoShare > 0, 1L, 0L)) %>%
  group_by(DayOfWeek, PromoDay) %>%
  summarise(
    mean_sales = mean(Sales, na.rm = TRUE),
    sd_sales = sd(Sales, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

promo_summary

promo_lm <- agg %>%
  filter(Sales > 0) %>%
  mutate(
    DayOfWeek = factor(DayOfWeek),
    Month = factor(Month)
  ) %>%
  lm(log1p(Sales) ~ PromoShare + OpenStores + DayOfWeek + Month, data = .)

coef(summary(promo_lm))["PromoShare", ]
```

## Day-of-Week Effect

```{r day-of-week-effect, fig.width=8, fig.height=4}
dow_summary <- agg %>%
  filter(Sales > 0) %>%
  group_by(DayOfWeek) %>%
  summarise(mean_sales = mean(Sales, na.rm = TRUE), .groups = "drop")

ggplot(dow_summary, aes(x = factor(DayOfWeek), y = mean_sales)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Average Sales by Day Of Week - Store Type A",
    x = "DayOfWeek (1 = Monday, 7 = Sunday)",
    y = "Mean Sales"
  )
```

## ACF and PACF

```{r acf-pacf, fig.width=11, fig.height=4}
par(mfrow = c(1, 2))
acf(agg$Sales, lag.max = 60, main = "ACF - Store Type A")
pacf(agg$Sales, lag.max = 60, main = "PACF - Store Type A")
par(mfrow = c(1, 1))
```

## STL Decomposition (Weekly Frequency)

```{r stl-decomposition, fig.width=11, fig.height=7}
sales_ts <- ts(agg$Sales, frequency = 7)
stl_fit <- stl(sales_ts, s.window = "periodic")
plot(stl_fit)
```

## Stationarity Tests

```{r stationarity-tests}
adf_result <- adf.test(agg$Sales)
kpss_result <- kpss.test(agg$Sales)

adf_result
kpss_result
```

## EDA Notes for Modeling

- Strong weekly behavior (period 7) supports seasonal ARIMA terms.
- Promo share, open stores, and day-of-week effects suggest useful external regressors for ARMAX.
- Stationarity test outcomes guide whether differencing is required.
