---
title: "Rossmann Sales Data Wrangling"
output:
  pdf_document: default
  html_document: default
date: "2026-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Data Wrangling

```{r}
library(tidyverse)
library(lubridate)
```

### First, let's load the data

```{r}
train <- read_csv("../data/raw/train.csv")
store <- read_csv("../data/raw/store.csv")
test  <- read_csv("../data/raw/test.csv")

head(train)
head(store)
```

### We focus on store model a, and aggregate this to form a time series
```{r}
# Join train.csv and store.csv
df_A <- train %>%
  left_join(store, by = "Store") %>%
  filter(StoreType == "a", Open == 1) %>%
  mutate(Date = as.Date(Date))

# =========================
# VERSION 1: CLEAN (UNIVARIATE TS)
# =========================
df_A_daily_clean <- df_A %>%
  group_by(Date) %>%
  summarise(
    Sales = sum(Sales),
    .groups = "drop"
  ) %>%
  arrange(Date) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Week = week(Date)
  )

head(df_A_daily_clean)

# =========================
# VERSION 2: WITH EXTERNAL VARIABLES
# =========================
df_A_daily_full <- df_A %>%
  group_by(Date) %>%
  summarise(
    Sales = sum(Sales),
    n_stores = n(),
    Promo = mean(Promo),
    SchoolHoliday = mean(SchoolHoliday),
    StateHoliday = mean(StateHoliday != "0"),
    .groups = "drop"
  ) %>%
  arrange(Date) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Week = week(Date)
  )

head(df_A_daily_full)

# =========================
# SAVE BOTH DATASETS
# =========================
write_csv(df_A_daily_clean, "../data/processed/daily_store_a_clean.csv")
write_csv(df_A_daily_full, "../data/processed/daily_store_a_full.csv")
```
In addition to sales, several time-varying explanatory variables were included. The Promo variable represents the proportion of stores running promotions on a given day, while SchoolHoliday and StateHoliday capture the proportion of stores affected by school and public holidays, respectively. These variables provide additional context for changes in sales beyond purely temporal patterns.

### Time series object
```{r}
ts_A <- ts(df_A_daily_clean$Sales, frequency = 7)
plot(ts_A)
```

The aggregated time series exhibits strong weekly seasonality, with noticeable drops in sales on certain days (e.g., Sundays), reflecting reduced store activity rather than true demand fluctuations. Notice that these drops in sales are due to most of the closed stores on certain days, but it is not entirely zero as some stores remain open that day.