---
title: "Rossmann Sales Data Wrangling"
output:
  pdf_document: default
  html_document: default
date: "2026-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Data Wrangling

```{r}
library(tidyverse)
library(lubridate)
```

### Load the data

```{r}
train <- read_csv("../data/raw/train.csv")
store <- read_csv("../data/raw/store.csv")
test  <- read_csv("../data/raw/test.csv")

head(train)
head(store)
```

### Build Store Type A base tables

```{r}
train_a_store <- train %>%
  left_join(store, by = "Store") %>%
  filter(StoreType == "a") %>%
  mutate(Date = as.Date(Date))

test_a_store <- test %>%
  left_join(store, by = "Store") %>%
  filter(StoreType == "a") %>%
  mutate(Date = as.Date(Date))

glimpse(train_a_store)
glimpse(test_a_store)
```

### Create train and test modeling panels for Store Type A

```{r}
train_a_model <- train_a_store %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    DayOfWeek = wday(Date, week_start = 1),
    IsoWeek = isoweek(Date),
    IsWeekend = if_else(DayOfWeek >= 6, 1L, 0L),
    Open = coalesce(Open, 1L),
    Promo = coalesce(Promo, 0L),
    SchoolHoliday = coalesce(SchoolHoliday, 0L)
  )

test_a_model <- test_a_store %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    DayOfWeek = wday(Date, week_start = 1),
    IsoWeek = isoweek(Date),
    IsWeekend = if_else(DayOfWeek >= 6, 1L, 0L),
    Open = coalesce(Open, 1L),
    Promo = coalesce(Promo, 0L),
    SchoolHoliday = coalesce(SchoolHoliday, 0L)
  )

glimpse(train_a_model)
glimpse(test_a_model)
```
In addition to sales, several time-varying explanatory variables were included. The Promo variable represents the proportion of stores running promotions on a given day, while SchoolHoliday and StateHoliday capture the proportion of stores affected by school and public holidays, respectively. These variables provide additional context for changes in sales beyond purely temporal patterns.

### Aggregate Store Type A to complete daily train and test series

```{r}
train_a_agg_daily <- train_a_model %>%
  group_by(Date) %>%
  summarise(
    Sales = sum(Sales, na.rm = TRUE),
    OpenStores = sum(Open == 1, na.rm = TRUE),
    PromoStores = sum(Promo == 1, na.rm = TRUE),
    SchoolHolidayShare = mean(SchoolHoliday, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    PromoShare = if_else(OpenStores > 0, PromoStores / OpenStores, 0)
  ) %>%
  complete(
    Date = seq(min(Date), max(Date), by = "day"),
    fill = list(
      Sales = 0,
      OpenStores = 0,
      PromoStores = 0,
      PromoShare = 0,
      SchoolHolidayShare = 0
    )
  ) %>%
  mutate(
    SchoolHolidayShare = replace_na(SchoolHolidayShare, 0),
    Year = year(Date),
    Month = month(Date),
    DayOfWeek = wday(Date, week_start = 1),
    IsoWeek = isoweek(Date),
    IsWeekend = if_else(DayOfWeek >= 6, 1L, 0L)
  ) %>%
  arrange(Date) %>%
  select(
    Date,
    Sales,
    OpenStores,
    PromoStores,
    PromoShare,
    SchoolHolidayShare,
    DayOfWeek,
    Year,
    Month,
    IsoWeek,
    IsWeekend
  )

test_a_agg_daily <- test_a_model %>%
  group_by(Date) %>%
  summarise(
    OpenStores = sum(Open == 1, na.rm = TRUE),
    PromoStores = sum(Promo == 1, na.rm = TRUE),
    SchoolHolidayShare = mean(SchoolHoliday, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    PromoShare = if_else(OpenStores > 0, PromoStores / OpenStores, 0)
  ) %>%
  complete(
    Date = seq(min(Date), max(Date), by = "day"),
    fill = list(
      OpenStores = 0,
      PromoStores = 0,
      PromoShare = 0,
      SchoolHolidayShare = 0
    )
  ) %>%
  mutate(
    SchoolHolidayShare = replace_na(SchoolHolidayShare, 0),
    Year = year(Date),
    Month = month(Date),
    DayOfWeek = wday(Date, week_start = 1),
    IsoWeek = isoweek(Date),
    IsWeekend = if_else(DayOfWeek >= 6, 1L, 0L)
  ) %>%
  arrange(Date) %>%
  select(
    Date,
    OpenStores,
    PromoStores,
    PromoShare,
    SchoolHolidayShare,
    DayOfWeek,
    Year,
    Month,
    IsoWeek,
    IsWeekend
  )

head(train_a_agg_daily)
head(test_a_agg_daily)
```

### Additional outputs (Version 1 and Version 2)

```{r}
# Version 1: clean univariate daily sales series
df_A_daily_clean <- train_a_model %>%
  group_by(Date) %>%
  summarise(
    Sales = sum(Sales, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Date) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Week = week(Date)
  )

# Version 2: daily series with external variables
df_A_daily_full <- train_a_model %>%
  group_by(Date) %>%
  summarise(
    Sales = sum(Sales, na.rm = TRUE),
    n_stores = n(),
    Promo = mean(Promo, na.rm = TRUE),
    SchoolHoliday = mean(SchoolHoliday, na.rm = TRUE),
    StateHoliday = mean(if_else(is.na(StateHoliday) | StateHoliday == "0", 0, 1), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Date) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Week = week(Date)
  )

head(df_A_daily_clean)
head(df_A_daily_full)
```

### Sanity checks

```{r}
date_span <- as.integer(max(train_a_agg_daily$Date) - min(train_a_agg_daily$Date)) + 1
missing_days <- date_span - nrow(train_a_agg_daily)

tibble(
  n_store_type_a_train_stores = n_distinct(train_a_model$Store),
  n_store_type_a_test_stores = n_distinct(test_a_model$Store),
  n_train_rows = nrow(train_a_model),
  n_test_rows = nrow(test_a_model),
  n_train_agg_daily_rows = nrow(train_a_agg_daily),
  n_test_agg_daily_rows = nrow(test_a_agg_daily),
  missing_days_in_agg = missing_days
)
```

### Save processed outputs

```{r}
dir.create("../data/processed", showWarnings = FALSE, recursive = TRUE)

write_csv(train_a_model, "../data/processed/store_type_a_train_store_daily.csv")
write_csv(test_a_model, "../data/processed/store_type_a_test_store_daily.csv")
write_csv(train_a_agg_daily, "../data/processed/store_type_a_train_aggregate_daily.csv")
write_csv(test_a_agg_daily, "../data/processed/store_type_a_test_aggregate_daily.csv")
write_csv(df_A_daily_clean, "../data/processed/daily_store_a_clean.csv")
write_csv(df_A_daily_full, "../data/processed/daily_store_a_full.csv")
```
