---
title: "4.4 ARMAX Method: Aggregate-Daily (Store Type A)"
output:
  pdf_document: default
  html_document: default
date: "2026-04-09"
---

## Section Overview

**Goal**
By building and comparing various ARMAX models for aggregated daily sales of Store Type A, we wish to answer the following question by the end of this section:

*Can we improve sales forecasts by using external variables, and which predictor combination works best?*

This notebook:

1. Reads aggregate processed files directly.
2. Uses a single train/holdout split.
3. Selects the best ARMAX specification by train AIC.
4. Evaluates the selected model on holdout using fixed-coefficient 1-step forecasts.
5. Produces aggregate-daily test forecasts.

Summary of Approach: 

4.4.1 Define forecasting assumptions and available information
4.4.2 Specify candidate ARMAX models with various specifications
4.4.3 Split data into training and holdout sets (merge wrangling later?)
4.4.4 Explore variable relationships using CCF diagnostics
4.4.5 Fit ARMAX models and select the best specification
4.4.6 Evaluate forecasting performance on the holdout set
4.4.7 Perform residual diagnostics
4.4.8 Refit the best model on full training data and forecast the test period

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 4.4.0 Data Preparation and Feature Construction
\textcolor{red}{Shall we move library set ups to the beginning of the report later?}

```{r libraries}
library(tidyverse)
library(forecast)

seasonal_frequency <- 7L

train_data <- readr::read_csv(
  "../../data/processed/store_type_a_train_aggregate_daily.csv",
  show_col_types = FALSE
) |>
  dplyr::mutate(
    Date = as.Date(Date),
    PromoShare = dplyr::if_else(OpenStores > 0, PromoStores / OpenStores, 0),
    DayOfWeek = as.integer(DayOfWeek),
    Month = as.integer(Month)
  ) |>
  dplyr::arrange(Date)

test_data <- readr::read_csv(
  "../../data/processed/store_type_a_test_aggregate_daily.csv",
  show_col_types = FALSE
) |>
  dplyr::mutate(
    Date = as.Date(Date),
    PromoShare = dplyr::if_else(OpenStores > 0, PromoStores / OpenStores, 0),
    DayOfWeek = as.integer(DayOfWeek),
    Month = as.integer(Month)
  ) |>
  dplyr::arrange(Date)

head(train_data)
```

## 4.4.1 Forecasting Assumption and Information Set

This ARMAX setup is a conditional forecast: we assume day-$t$ inputs are known when forecasting day $t$.

For a 1-step-ahead forecast at day $t$, the model uses:

$$
\mathcal{F}_{t-1} \cup \{X_t^{planned}\}
$$

Here, $X_t$ includes `PromoShare`, `OpenStores`, and calendar covariates.

Practical interpretation:

1. `DayOfWeek` and `Month` are deterministic and known in advance.
2. `PromoShare` and `OpenStores` are valid only if planned and available by the end of day $t-1$.
3. If you use realized day-$t$ values, holdout RMSE may be overly optimistic due to information leakage beyond $\mathcal{F}_{t-1}$.

\textcolor{red}{All makes great sense - do we need to prove this is the case in the actual dataset when we practice actual analysis?}

## 4.4.2 Candidate ARMAX Model Specifications

```{r model-specs}
model_specs <- list(
  ARMAX_Promo_Open = stats::as.formula("~ PromoShare + OpenStores"),
  ARMAX_Promo_Open_DOW = stats::as.formula("~ PromoShare + OpenStores + DayOfWeek"),
  ARMAX_Promo_Open_DOW_Month = stats::as.formula("~ PromoShare + OpenStores + DayOfWeek + Month")
)

model_specs
```

## 4.4.3 Holdout Split Setup
\textcolor{red}{potentially merge with wrangling for final version}

```{r holdout-split}
set.seed(443)

holdout_prop <- 0.20
n_total <- nrow(train_data)
n_train <- floor((1 - holdout_prop) * n_total)

train_subset <- train_data |> dplyr::slice(1:n_train)
holdout_subset <- train_data |> dplyr::slice((n_train + 1):n_total)

holdout_split <- tibble::tibble(
  n_total = n_total,
  n_train = n_train,
  n_holdout = n_total - n_train,
  train_start = min(train_subset$Date),
  train_end = max(train_subset$Date),
  holdout_start = min(holdout_subset$Date),
  holdout_end = max(holdout_subset$Date)
)

holdout_split
```

## 4.4.4 Diagnostic Analysis: Cross-Correlation (CCF)

```{r ccf-diagnostics, fig.width=10, fig.height=4}
diag_data <- train_subset |>
  dplyr::transmute(
    Sales,
    PromoShare,
    OpenStores
  ) |>
  dplyr::mutate(
    dSales = c(NA_real_, diff(Sales)),
    dPromoShare = c(NA_real_, diff(PromoShare)),
    dOpenStores = c(NA_real_, diff(OpenStores))
  ) |>
  tidyr::drop_na()

old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))

stats::ccf(
  diag_data$dPromoShare,
  diag_data$dSales,
  lag.max = 21,
  main = "CCF: dPromoShare vs dSales"
)

stats::ccf(
  diag_data$dOpenStores,
  diag_data$dSales,
  lag.max = 21,
  main = "CCF: dOpenStores vs dSales"
)

par(old_par)
```

## 4.4.5 ARMAX Modeling: Function Setup, Fitting and Selection

```{r armax-functions}
build_xreg <- function(df, model_formula) {
  df <- df |>
    dplyr::mutate(
      DayOfWeek = factor(DayOfWeek, levels = 1:7),
      Month = factor(Month, levels = 1:12)
    )

  mm <- stats::model.matrix(model_formula, data = df)
  mm[, colnames(mm) != "(Intercept)", drop = FALSE]
}

rmse_vec <- function(actual, pred) {
  sqrt(mean((actual - pred) ^ 2, na.rm = TRUE))
}

clean_xreg <- function(xreg_train, xreg_valid) {
  xreg_train <- as.matrix(xreg_train)
  xreg_valid <- as.matrix(xreg_valid)

  keep_non_constant <- apply(xreg_train, 2, function(x) stats::var(x, na.rm = TRUE) > 0)
  xreg_train <- xreg_train[, keep_non_constant, drop = FALSE]
  xreg_valid <- xreg_valid[, keep_non_constant, drop = FALSE]

  qr_fit <- qr(xreg_train)
  keep_idx <- sort(qr_fit$pivot[seq_len(qr_fit$rank)])
  xreg_train <- xreg_train[, keep_idx, drop = FALSE]
  xreg_valid <- xreg_valid[, keep_idx, drop = FALSE]

  list(train = xreg_train, valid = xreg_valid)
}

prepare_xreg_for_split <- function(data, model_formula, n_train) {
  xreg_full_raw <- build_xreg(data, model_formula)
  xreg_train_raw <- xreg_full_raw[seq_len(n_train), , drop = FALSE]

  xreg_clean <- clean_xreg(xreg_train_raw, xreg_full_raw)

  list(
    train = xreg_clean$train,
    full = xreg_clean$valid
  )
}

fit_armax_train <- function(data, model_formula, n_train) {
  y_train <- ts(data$Sales[seq_len(n_train)], frequency = seasonal_frequency)
  xreg_obj <- prepare_xreg_for_split(data, model_formula, n_train)

  fit <- forecast::auto.arima(
    y_train,
    xreg = xreg_obj$train,
    seasonal = TRUE,
    allowmean = FALSE,
    allowdrift = FALSE,
    stepwise = TRUE,
    approximation = TRUE
  )

  list(
    fit = fit,
    xreg_full = xreg_obj$full
  )
}

resolve_seasonal_period <- function(ord, y_ts, fallback_period = 7L) {
  m_raw <- suppressWarnings(as.integer(unname(ord["m"])))
  freq_raw <- suppressWarnings(as.integer(stats::frequency(y_ts)))

  if (!is.na(m_raw) && m_raw > 1L) {
    return(m_raw)
  }

  if (!is.na(freq_raw) && freq_raw > 1L) {
    return(freq_raw)
  }

  fallback_period
}

extract_arima_spec <- function(train_fit_obj, y_ts) {
  ord <- forecast::arimaorder(train_fit_obj)

  list(
    ord = ord,
    order = c(
      as.integer(unname(ord["p"])),
      as.integer(unname(ord["d"])),
      as.integer(unname(ord["q"]))
    ),
    seasonal_order = c(
      as.integer(unname(ord["P"])),
      as.integer(unname(ord["D"])),
      as.integer(unname(ord["Q"]))
    ),
    seasonal_period_raw = unname(ord["m"]),
    seasonal_period_used = resolve_seasonal_period(ord, y_ts)
  )
}

fixed_coef_holdout_forecast <- function(data, n_train, train_fit_obj, xreg_full) {
  y_all <- ts(data$Sales, frequency = seasonal_frequency)
  spec <- extract_arima_spec(train_fit_obj, y_all)

  fixed_fit <- stats::arima(
    y_all,
    order = spec$order,
    seasonal = list(
      order = spec$seasonal_order,
      period = spec$seasonal_period_used
    ),
    xreg = xreg_full,
    include.mean = FALSE,
    fixed = as.numeric(train_fit_obj$coef),
    transform.pars = FALSE,
    optim.control = list(maxit = 0),
    method = "ML"
  )

  fc_all <- as.numeric(y_all - fixed_fit$residuals)
  holdout_idx <- (n_train + 1):length(y_all)

  tibble::tibble(
    Date = data$Date[holdout_idx],
    actual = as.numeric(y_all[holdout_idx]),
    predicted = fc_all[holdout_idx]
  )
}
```

### Select ARMAX Specification by Train AIC
\textcolor{red}{Yolanda may add RMSE here in a bit (and recite the numbers in the RMSE section for overall comparison) - depends on where the RMSE calculations make most sense}

```{r train-aic-selection}
y_train_for_period <- ts(train_data$Sales[seq_len(n_train)], frequency = seasonal_frequency)

train_fit_registry <- purrr::imap(model_specs, function(model_formula, model_name) {
  fit_armax_train(
    data = train_data,
    model_formula = model_formula,
    n_train = n_train
  )
})

aic_table <- purrr::imap_dfr(train_fit_registry, function(obj, model_name) {
  spec <- extract_arima_spec(obj$fit, y_train_for_period)
  seasonal_period_label <- as.character(as.integer(spec$seasonal_period_used))

  tibble::tibble(
    model = model_name,
    train_aic = stats::AIC(obj$fit),
    seasonal_period_raw = spec$seasonal_period_raw,
    seasonal_period_used = spec$seasonal_period_used,
    arima_order = sprintf(
      "(%d,%d,%d)(%d,%d,%d)[%s]",
      spec$order[1],
      spec$order[2],
      spec$order[3],
      spec$seasonal_order[1],
      spec$seasonal_order[2],
      spec$seasonal_order[3],
      seasonal_period_label
    )
  )
}) |>
  dplyr::arrange(train_aic)

aic_table

seasonal_period_issues <- aic_table |>
  dplyr::filter(is.na(seasonal_period_raw))

if (nrow(seasonal_period_issues) > 0) {
  warning("At least one ARMAX candidate reports NA seasonal period from arimaorder(). Using seasonal_period_used fallback.")
}

seasonal_period_issues
```

## 4.4.6 Fixed-Coefficient 1-Step Holdout Evaluation

```{r holdout-evaluation}
best_model <- aic_table |>
  dplyr::slice_min(order_by = train_aic, n = 1)

best_model_name <- best_model$model[[1]]
best_model_formula <- model_specs[[best_model_name]]
best_train_fit <- train_fit_registry[[best_model_name]]$fit
best_xreg_full <- train_fit_registry[[best_model_name]]$xreg_full

holdout_predictions <- fixed_coef_holdout_forecast(
  data = train_data,
  n_train = n_train,
  train_fit_obj = best_train_fit,
  xreg_full = best_xreg_full
)

armax_holdout_summary <- tibble::tibble(
  model = best_model_name,
  train_aic = best_model$train_aic[[1]],
  holdout_rmse = rmse_vec(holdout_predictions$actual, holdout_predictions$predicted),
  n_holdout = nrow(holdout_predictions)
)

armax_holdout_summary
```

## 4.4.7 Residual Diagnostics

```{r residual-diagnostics}
forecast::checkresiduals(best_train_fit)

residual_diagnostics <- tibble::tibble(
  model = best_model_name,
  sigma2 = best_train_fit$sigma2,
  ljung_box_pvalue = stats::Box.test(
    residuals(best_train_fit),
    lag = 14,
    type = "Ljung-Box",
    fitdf = sum(best_train_fit$arma[1:4])
  )$p.value
)

residual_diagnostics
```

```{r holdout-plot, fig.width=9, fig.height=4}
ggplot(holdout_predictions, aes(x = Date)) +
  geom_line(aes(y = actual, color = "Actual")) +
  geom_line(aes(y = predicted, color = "Predicted")) +
  labs(
    title = "ARMAX Holdout: Fixed-Coefficient 1-Step Forecasts",
    x = "Date",
    y = "Sales",
    color = "Series"
  )
```

### Best-Performing ARMAX Model

```{r choose-best}
best_model

best_model_name
best_model_formula
```

## 4.4.8 Test Forecasting with Final Model 

```{r full-fit-and-forecast}
y_full <- ts(train_data$Sales, frequency = seasonal_frequency)

xreg_train_full_raw <- build_xreg(train_data, best_model_formula)
xreg_test_raw <- build_xreg(test_data, best_model_formula)

xreg_clean_full <- clean_xreg(xreg_train_full_raw, xreg_test_raw)
xreg_train_full <- xreg_clean_full$train
xreg_test <- xreg_clean_full$valid

best_fit <- forecast::auto.arima(
  y_full,
  xreg = xreg_train_full,
  seasonal = TRUE,
  allowmean = FALSE,
  allowdrift = FALSE,
  stepwise = TRUE,
  approximation = TRUE
)
test_forecast <- as.numeric(forecast::forecast(best_fit, xreg = xreg_test, h = nrow(test_data))$mean)

aggregate_test_forecast <- test_data |>
  dplyr::transmute(
    Date,
    predicted_aggregate_sales = pmax(0, test_forecast)
  )

head(aggregate_test_forecast)
```

## Save Outputs

```{r save-results}
dir.create("results", showWarnings = FALSE, recursive = TRUE)

readr::write_csv(holdout_split, "results/armax_holdout_split.csv")
readr::write_csv(aic_table, "results/armax_aic_table.csv")
readr::write_csv(holdout_predictions, "results/armax_holdout_predictions.csv")
readr::write_csv(armax_holdout_summary, "results/armax_holdout_rmse_summary.csv")
readr::write_csv(aggregate_test_forecast, "results/armax_aggregate_test_forecast.csv")
```
