## 4.3 ARIMA Method

### Model Identification

```{r}
sales_ts <- ts(agg$Sales, frequency = 7)
auto_model <- auto.arima(sales_ts)
summary(auto_model)
```

Automated model selection using auto_model identified **ARIMA(4,0,2)(0,1,1)[7]** as the initial candidate based on AIC minimization.  

It indicates that there's: 
- No non-seasonal differencing ($d=0$), consistent with prior stationarity tests  
- One seasonal difference ($D=1$) to account for weekly patterns  
- Significant autoregressive structure at lower lags, with diminishing contributions at higher lags  

Also, the coeffieicnet estimates suggest that the first two autoregressive terms contribute most strongly, while higher-order terms don't have much of an impact. 

Training performance metrics indicate reasonable fit, with pretty low residual autocorrelation (ACF1 ≈ 0) and MASE < 1. 

---

### Residuals

```{r}
checkresiduals(auto_model)
```
We check if the model left any patterns behind in the residuals. If residuals look like random noise, the model did a good job.

When we ran checkresiduals, we found that while errors are approximately centered and have a near-normal behavior, there is still some indicators of remaining autocorrelation. Residual ACF shows spikes at seasonal lags (7 and 21), and the Ljung-Box test is statistically significant (p < 0.05) which indicaed that residuals are not fully white noise  This suggests that the baseline model does not fully capture the underlying seasonal structure.

**Top plot**: residuals bounce randomly around zero, mostly good, but some large spikes around early 2014 which match the holiday outliers we saw in EDA
**ACF plot**: a few spikes outside the blue lines, especially around lags 7 and 21, indicating the model hasn't fully captured the weekly pattern
**Histogram**: roughly bell shaped and centered at zero, meaning errors are fairly normally distributed with no consistent bias in one direction
**Ljung-Box test result**: p-value = 0.0003958 (less than 0.05) → there is still some remaining pattern in the residuals, which tells us the model could potentially be improved and suggests we can perhaps try alternative models.

---

### Candidate Models 

The best AIC model isn't always the best forecasting model. We try variations motivated by our EDA, specifically p=1 and p=2 because our PACF showed spikes at lags 7 and 14. Here are the several alternative candidate modesls we considered:

```{r}
model1 <- auto_model
model2 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(0,1,2))
model3 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(1,1,1))
model4 <- Arima(sales_ts, order=c(2,0,2), seasonal=c(0,1,1))
model5 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(2,1,1))
```

The reasons we're trying these different models are: 
- **Model2**:  ACF of residuals showed spikes at seasonal lags, so we tried increasing Q
- **Model3**: PACF showed spike at lag 7, suggesting a seasonal AR term P=1 might help
- **Model4**: ar3 and ar4 coefficients were close to zero, suggesting p=2 might be enough
- **Model5**: PACF showed spike at lag 14 as well, suggesting P=2 might capture that

---

### Models Comparison

```{r}
AIC(model1, model2, model3, model4, model5)
```

ARIMA(4,0,2)(0,1,1)[7]: 28,077.52 (lowest AIC)
ARIMA(4,0,2)(2,1,1)[7]: 28,080.97
ARIMA(2,0,2)(0,1,1)[7]: 28,087.45
ARIMA(4,0,2)(1,1,1)[7]: 28,088.42
ARIMA(4,0,2)(0,1,2)[7]: 28,103.65

The original model, **ARIMA(4,0,2)(0,1,1)[7]** has the lowest AIC. However, given there is residual autocorrelation, best AIC doesn't guarantee best out-of-sample forecasts and accuracy. All candidate models are should still be carried out for RMSE comparison.