## 4.3 ARIMA Method

### Model Identification

```{r}
sales_ts <- ts(agg$Sales, frequency = 7)
auto_model <- auto.arima(sales_ts)
summary(auto_model)
```

The automated model selection procedure identified **ARIMA(4,0,2)(0,1,1)[7]** as the initial candidate based on AIC minimization.  

The specification indicates:
- No non-seasonal differencing ($d=0$), consistent with prior stationarity tests  
- One seasonal difference ($D=1$) to account for weekly patterns  
- Significant autoregressive structure at lower lags, with diminishing contributions at higher lags  

Coefficient estimates suggest that the first two autoregressive terms contribute most strongly, while higher-order terms are relatively small.

Training performance metrics indicate reasonable fit, with low residual autocorrelation (ACF1 ≈ 0) and MASE < 1, suggesting improvement over a naive benchmark.

---

### Residual Diagnostics

```{r}
checkresiduals(auto_model)
```

Residual diagnostics indicate that while errors are approximately centered and exhibit near-normal behavior, there is evidence of remaining autocorrelation.  

In particular:
- Residual ACF shows spikes at seasonal lags (7 and 21)  
- The Ljung-Box test is statistically significant (p < 0.05), indicating that residuals are not fully white noise  

This suggests that the baseline model does not fully capture the underlying seasonal structure.

---

### Candidate Model Specification

Based on residual diagnostics and coefficient significance, several alternative specifications were considered:

```{r}
model1 <- auto_model
model2 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(0,1,2))
model3 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(1,1,1))
model4 <- Arima(sales_ts, order=c(2,0,2), seasonal=c(0,1,1))
model5 <- Arima(sales_ts, order=c(4,0,2), seasonal=c(2,1,1))
```

These variations were motivated by:
- Increasing seasonal MA terms to address residual autocorrelation  
- Introducing seasonal AR components based on PACF behavior  
- Reducing non-seasonal AR order due to weak higher-order coefficients  

---

### Model Comparison

```{r}
AIC(model1, model2, model3, model4, model5)
```

| Model | AIC |
|------|-----|
| ARIMA(4,0,2)(0,1,1)[7] | 28077.52 |
| ARIMA(4,0,2)(2,1,1)[7] | 28080.97 |
| ARIMA(2,0,2)(0,1,1)[7] | 28087.45 |
| ARIMA(4,0,2)(1,1,1)[7] | 28088.42 |
| ARIMA(4,0,2)(0,1,2)[7] | 28103.65 |

The original specification, **ARIMA(4,0,2)(0,1,1)[7]**, achieved the lowest AIC and is therefore selected as the preferred model within this candidate set.  

However, given the presence of residual autocorrelation, all candidate models are retained for further evaluation using out-of-sample forecast accuracy in the subsequent section.